### 2022-08-08 letterhead.tex

\setlength\topmargin{-0.50 in}
{\normalfont
\begin{flushright}
{\Large {\bf \color{Maroon} University of Maryland School of Pharmacy}} 
   %\hspace{\fill} 
%\includegraphics[height=20truept]{OUlogo}
\end{flushright}
\vspace{-0.5cm}
\begin{tabular*}{\linewidth}{c} \hline \end{tabular*}
\begin{flushright}
{\footnotesize {\bf {\em Jana Shen, Ph.D.}} \hspace{\fill} 
        {\em  Department of Pharmaceutical Sciences} }     \\
{\footnotesize {\em Professor} \hspace{\fill} 
        {\em  Computer-Aided Drug Design Center} }                    \\
{\footnotesize {\em Codirector, Computer-Aided Drug Design Center} \hspace{\fill}
        {\em University of Maryland School of Pharmacy} }  \\
{\footnotesize {\em Phone: (410) 706-4187} \hspace{\fill} 
        {\em Baltimore, MD 21201} }          \linebreak
{\footnotesize {\em E-mail: \url{jana.shen@rx.umaryland.edu}} \hspace{\fill} 
        {\em \url{http://www.computchem.org}} }                               \\
        \end{flushright}
\vspace{-0.5cm}
}